Statistical Inference for inter-arrival times of extreme events in bursty time series

### 2020-08-01 article_springer/article_springer.Rmd
bügele das Englisch

  A new model for the _inter-exceedance times_ of events 
  above high thresholds is proposed. For high thresholds and 
  infinite-mean waiting times, it is shown that the times between threshold 
  crossings are Mittag-Leffler distributed, and thus form a "fractional 
  Poisson Process" which generalizes the standard Poisson Process of threshold
  exceedances.
  Graphical means of estimating model parameters and assessing 
  model fit are provided. Along the way, the inference method gets applied to a real-world
  bursty time series, and it is shown how the memory of the Mittag-Leffler distribution affects 
  the predictive distribution for the time until the next extreme event.

@@

For inter-exceedance *times* (IETs), the same idea applies. 
By the Theorem, the inter-exceedance times converge to a Mittag-Leffler
distribution (MLD), and importantly, the tail parameter $\beta$ is 
*independent* of (or *stable* with respect to) the choice of threshold.
For high thresholds, the IET distribution may fit an MLD well, but few
IETs may be left for a fit (high variance). 
For low thresholds, their distribution may be not well represented
by an MLD; or, too many events may be interpreted as 
shocks, resulting in a biased estimate of the parameters of the limiting MLD.
<!-- For illustration, consider the case where measurements are taken daily: -->
<!-- Then at the lowest threshold, all IETs equal one day, -->
<!-- leaving no hint at a power-law distribution whatsoever. Raising the threshold -->
<!-- may or may not produce IETs that start exhibiting a power-law; and both cases -->
<!-- are represented by the MLD, with $\beta < 1$ or $\beta = 1$. -->

<!-- Suppose the "correct" threshold $\ell_0$ was known, in the sense that  -->
<!-- thresholding at $\ell_0$ would result precisely in an underlying MRP  -->
<!-- $\{(W_i, J_i): i \in I\}$ as defined in Section \ref{sec:CTRE}.  -->
In case we would have data of a MRP $\{(W_i, J_i): i \in I\}$ as defined in Section \ref{sec:CTRE}, we could apply any tail (and scale) estimator such as e.g. Hill's estimator [@hill1975simple] to the $W_i$ and thus infer the distribution of inter-exceedance times. However, in real-world datasets, the MRP assumption is too strong for low thresholds. The unknown data generating process is likely to be more complex, with dependencies or mutliple data generating mechanisms at low thresholds, structures that possibly vanish for higher thresholds. Fitting only the IETs means that the iid assumption applies to exceedances and IETs only.

<!-- But guessing the correct threshold $\ell_0$ and proceeding with this -->
<!-- type of inference is less favourable to modelling IETs with the MLD,  -->
<!-- as we now explain. -->
<!-- In real-world datasets, a "correct" threshold may not exist: the data  -->
<!-- $\{(T_i, X_i): i \in I\}$ may be a mixture of an MRP with some other -->
<!-- time series $\{(T'_i, X'_i): i \in I'\}$, and the supports $\mathcal S'$ -->
<!-- and $\mathcal S$ of the distributions -->
<!-- of $X'_i$ and the magnitudes $J_i$ (aka the MRP marks) may overlap. -->
<!-- Then thresholding at e.g. $\ell_0 = \sup \mathcal S'$  -->
<!-- no longer results in a "clean" MRP: some $J_i$'s will be removed, meaning -->
<!-- that some $W_i$'s will be merged into a longer IET. -->
<!-- The more the threshold $\ell_0$ cuts into $\mathcal S$ from below, the more -->
<!-- the distributions of IETs will resemble a MLD, according to the Theorem. -->
<!-- As we do not know to what extent this effect occurs,  -->
<!-- and since the limit is an MLD, it is practicable to fit an MLD to the IETs to begin with. -->

>>>>

  A new model for the _inter-exceedance times_ of events 
  above high thresholds is proposed. For high thresholds and 
  infinite-mean waiting times, it is shown that the times between threshold 
  crossings are Mittag-Leffler distributed, and thus form a "fractional 
  Poisson Process" which generalizes the standard Poisson Process of threshold
  exceedances.
  Graphical means of estimating model parameters and assessing 
  model fit are provided. 
  The inference method is
  applied to an empirical bursty time series, and it is shown how the memory of the
  Mittag-Leffler distribution affects prediction of the time until
  the next extreme event."

@@

For inter-exceedance *times* (IETs), the same idea applies. 
By the Theorem, the inter-exceedance times converge to a Mittag-Leffler
distribution (MLD), and importantly, the tail parameter $\beta$ is 
*independent* of (or *stable* with respect to) the choice of threshold.
For high thresholds, the IET distribution may fit an MLD well, but few
IETs may be left for a fit (high variance). 
For low thresholds, their distribution may be not well represented
by an MLD; or, too many events may be interpreted as 
shocks, resulting in a biased estimate of the parameters of the limiting MLD.
If we had MRP data $\{(W_i, J_i): i \in I\}$ as defined in Section 
\ref{sec:CTRE}, we could apply any tail (and scale) estimator such as e.g. Hill's 
estimator [@hill1975simple] to the $W_i$ and thus infer the distribution of 
inter-exceedance times. However, in real-world datasets, the MRP assumption is too 
strong for low thresholds. The unknown data generating process is likely more 
complex, with dependencies or multiple data generating mechanisms at low thresholds 
and structures that may vanish for higher thresholds. Fitting only the IETs means 
that the iid assumption applies to exceedances and IETs only.
